Online methods

### 2013-06-27 docs/manuscript/methods.tex

Gene expression profiles were generated from whole blood collected with PAXgene TM tubes (QIAGEN, Valencia, CA) using Illumina HT12-v4.0 bead arrays. The Illumina HT-12 v4.0 chip contains 47,323 probes, although some probes are not assigned to RefSeq genes. We removed any probes that did not match the following criteria: contained a SNP within the probe sequence with MAF $ > 0.05$ within 1000 genomes data; did not map to a listed RefSeq gene; were not significantly expressed (based on a detection $p$-value $< 0.05$) in at least 90\% of samples. After this stringent QC 7339 probes remained for 2D-eQTL mapping.

@@

\paragraph{Fehrmann: $n=1240$}
The Fehrmann dataset \cite{Fehrmann2011} consists of whole peripheral blood samples of 1240 unrelated individuals from the United Kingdom and the Netherlands. Some of these individuals are patients, while others are healthy controls. Individuals were genotyped using the Illumina 610 Quad platform. RNA levels were quantified using the HT12v3 platform.

\paragraph{EGCUT: $n=891$}
The Estonian Genome Center of the University of Tartu (EGCUT) study \cite{Metspalu2004} consists of whole blood samples of 891 unrelated individuals from Estonia. They were genotyped using the Illumina 610 Quad platform. RNA levels were quantified using the HT12v3 platform.

\paragraph{CDHWB: $n=139$}
The Center for Health Discovery and Well Being (CDHWB) Study \cite{Preininger2013} is a population based cohort consisting of 139 individuals of European descent collected in Atlanta USA. Gene expression profiles were generated with Illumina HT-12 V3.0 arrays from whole blood collected from Tempus tubes that preserve RNA. Whole genome genotypes were measured using Illumina OmniQuad arrays. Due to the small sample size, most SNP pairs did not pass filtering in this dataset (185 remained) and so we have excluded it from the rest of the analysis.

>>>>

Gene expression profiles were generated from peripheral blood collected with PAXgene TM tubes (QIAGEN, Valencia, CA) using Illumina HT12-v4.0 bead arrays. The Illumina HT-12 v4.0 chip contains 47,323 probes, although some probes are not assigned to RefSeq genes. We removed any probes that did not match the following criteria: contained a SNP within the probe sequence with MAF $ > 0.05$ within 1000 genomes data; did not map to a listed RefSeq gene; were not significantly expressed (based on a detection $p$-value $< 0.05$) in at least 90\% of samples. After this stringent QC 7339 probes remained for 2D-eQTL mapping.

@@

\paragraph{Fehrmann: $n=1240$}
The Fehrmann dataset \cite{Fehrmann2011} consists of peripheral blood samples of 1240 unrelated individuals from the United Kingdom and the Netherlands. Some of these individuals are patients, while others are healthy controls. Individuals were genotyped using the Illumina 610 Quad platform. RNA levels were quantified using the HT12v3 platform.

\paragraph{EGCUT: $n=891$}
The Estonian Genome Center of the University of Tartu (EGCUT) study \cite{Metspalu2004} consists of peripheral blood samples of 891 unrelated individuals from Estonia. They were genotyped using the Illumina 610 Quad platform. RNA levels were quantified using the HT12v3 platform.

\paragraph{CDHWB: $n=139$}
The Center for Health Discovery and Well Being (CDHWB) Study \cite{Preininger2013} is a population based cohort consisting of 139 individuals of European descent collected in Atlanta USA. Gene expression profiles were generated with Illumina HT-12 V3.0 arrays from peripheral blood collected from Tempus tubes that preserve RNA. Whole genome genotypes were measured using Illumina OmniQuad arrays. Due to the small sample size, most SNP pairs did not pass filtering in this dataset (27 SNP pairs remained) and so we have excluded it from the rest of the analysis.
